---
title: "ST5188 PG01 Exploratory Data Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#| include: FALSE

# clear working directory
rm(list = ls())

scripts = list.files("../Code/Preprocessing", full.names = TRUE)

# source libraries and functions under `Code` files
for (file in scripts) {
  print(paste("Sourcing:", file))
  if (grepl("\\.R$", file)) {
    source(file)
  } else {
    print(paste("Python sourcing:", file))
    reticulate::source_python(file)
  }
}

# check if all functions have been sourced
ls()
```

### {.tabset .tabset-fade .tabset-pills}
Let's perform exploratory data analysis on our dataset!

```{r}
#| include: FALSE

# ensure it's globally available
knitr::opts_chunk$set(echo = TRUE)

plot_boundary <- function(subzone) {
  
  border <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = T) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  return(border)
}
  
img <- "../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif"
r <- rast(img) |>
  project("EPSG:4326") 
```

#### Changi

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10

subzone <- "CHANGI"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `long_changi`:

```{r, results='hide'}
# this section is to pre-process CHANGI data
subzone <- "CHANGI"

# extract, impute and combine data from images - swarnali + amira
df_changi <- combine_img(subzone) # takes around 3-4 mins
rownames(df_changi) <- NULL

# aggregate temporally and obtain final data - naomi
long_changi <- st_aggregate(df_changi, subzone)[[1]]
wide_changi <- st_aggregate(df_changi, subzone)[[2]]
```


By using `long_changi`, we can see how the LST values changes over time:

```{r}
#| echo: FALSE
#| fig.width: 10

gif_path <- py$create_lst_animation(long_changi, subzone = "CHANGI", fps = 2)
image_animate(image_read(gif_path), loop = 0)
```

```{r}
#| echo: FALSE
#| fig.width: 10

# this part is to write eda

```


#### Jurong East

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10

subzone <- "JURONG EAST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `long_je`:

```{r, results='hide'}
# this section is to pre-process CHANGI data
subzone <- "JURONG EAST"

# extract, impute and combine data from images - swarnali + amira
df_je <- combine_img(subzone) # takes around 3-4 mins
rownames(df_je) <- NULL

# aggregate temporally and obtain final data - naomi
long_je <- st_aggregate(df_je, subzone)[[1]]
wide_je <- st_aggregate(df_je, subzone)[[2]]
```

By using `long_je`, we can see how the LST values changes over time:

```{r}
#| echo: FALSE
#| fig.width: 10

gif_path <- py$create_lst_animation(long_je, subzone = "JURONG EAST", fps = 2)
include_graphics(gif_path)
```

```{r}
#| echo: FALSE
#| fig.width: 10

# this part is to write eda

```

#### Jurong West

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10

subzone <- "JURONG WEST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `long_jw`:

```{r, results='hide'}
# this section is to pre-process CHANGI data
subzone <- "JURONG WEST"

# extract, impute and combine data from images - swarnali + amira
df_jw <- combine_img(subzone) # takes around 3-4 mins
rownames(df_jw) <- NULL

# aggregate temporally and obtain final data - naomi
long_jw <- st_aggregate(df_jw, subzone)[[1]]
wide_jw <- st_aggregate(df_jw, subzone)[[2]]
```

By using `long_je`, we can see how the LST values changes over time:

```{r}
#| echo: FALSE
#| fig.width: 10

gif_path <- py$create_lst_animation(long_jw, subzone = "JURONG EAST", fps = 2)
image_animate(image_read(gif_path), loop = 0)
```

```{r}
#| echo: FALSE
#| fig.width: 10

# this part is to write eda

```