---
title: "ST5188 PG01 Exploratory Data Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#| include: FALSE
# this chunk is to source all our functions —————

# clear working directory
rm(list = ls())

scripts = list.files("../Code/Preprocessing", full.names = TRUE)

# source libraries and functions under `Code` files
for (file in scripts) {
  print(paste("Sourcing:", file))
  if (grepl("\\.R$", file)) {
    source(file)
  } else {
    print(paste("Python sourcing:", file))
    reticulate::source_python(file)
  }
}

# check if all functions have been sourced
ls()
```

### {.tabset .tabset-fade .tabset-pills}
Let's perform exploratory data analysis on our dataset!

```{r}
#| include: FALSE
# this chunk is to create the plot_boundary function and import img —————

plot_boundary <- function(subzone) {
  
  border <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = T) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  return(border)
}
  
img <- "../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif"
r <- rast(img) |>
  project("EPSG:4326") 
```

#### Changi

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot changi —————

subzone <- "CHANGI"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_changi_long`:

```{r, results='hide'}
# this chunk is to pre-process changi's landsat7 & landsat8 data —————

subzone <- "CHANGI"

# landsat 7
# extract, impute and combine data from images
df_changi7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
rownames(df_changi7) <- NULL

# landsat 8
df_changi8 <- combine_img(subzone, landsat_no = 8)
rownames(df_changi8) <- NULL

merged_changi <- merge_ldst(df_changi7, df_changi8)

# aggregate temporally and obtain final data
final_changi_long <- st_aggregate(merged_changi, subzone)[[1]]
final_changi_wide <- st_aggregate(merged_changi, subzone)[[2]]
```

By using `final_changi_long`, we can obtain the following EDA results:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

eda(final_changi_long)

gif_changi <- py$create_lst_animation(final_changi_long, subzone = "CHANGI", fps = 2)
include_graphics(gif_changi)
```

#### Jurong East

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot jurong east —————

subzone <- "JURONG EAST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_je_long`:

```{r, results='hide'}
# this chunk is to pre-process jurong east's landsat7 & landsat8 data —————

subzone <- "JURONG EAST"

# landsat 7
# extract, impute and combine data from images
df_je7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
rownames(df_je7) <- NULL

# landsat 8
df_je8 <- combine_img(subzone, landsat_no = 8)
rownames(df_je8) <- NULL

merged_je <- merge_ldst(df_je7, df_je8)

# aggregate temporally and obtain final data
final_je_long <- st_aggregate(merged_je, subzone)[[1]]
final_je_wide <- st_aggregate(merged_je, subzone)[[2]]
```

By using `final_je_long`, we can obtain the following EDA results:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

eda(final_je_long)

gif_je <- py$create_lst_animation(final_je_long, subzone = "JURONG EAST", fps = 2)
include_graphics(gif_je)
```

#### Jurong West

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot jurong west —————

subzone <- "JURONG WEST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_jw_long`:

```{r, results='hide'}
# this chunk is to pre-process jurong west's landsat7 & landsat8 data —————

subzone <- "JURONG WEST"

# landsat 7
# extract, impute and combine data from images
df_jw7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
rownames(df_jw7) <- NULL

# landsat 8
df_jw8 <- combine_img(subzone, landsat_no = 8)
rownames(df_jw8) <- NULL

merged_jw <- merge_ldst(df_jw7, df_jw8)

# aggregate temporally and obtain final data
final_jw_long <- st_aggregate(merged_jw, subzone)[[1]]
final_jw_wide <- st_aggregate(merged_jw, subzone)[[2]]
```

By using `final_jw_long`, we can see how the LST values changes over time:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

eda(final_jw_long)

gif_jw <- py$create_lst_animation(final_jw_long, subzone = "JURONG WEST", fps = 2)
include_graphics(gif_jw)
```
