---
title: "ST5188 PG01 Exploratory Data Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
#| include: FALSE
# this chunk is to source all our functions —————

# clear working directory
rm(list = ls())
reticulate::use_python("C:/Users/naomi/anaconda3/python.exe", required = TRUE) # run `which python3` in terminal to get your python path

scripts = list.files("../Code/Preprocessing", full.names = TRUE)

# source libraries and functions under `Code` files
for (file in scripts) {
  print(paste("Sourcing:", file))
  if (grepl("\\.R$", file)) {
    source(file)
  } else {
    print(paste("Python sourcing:", file))
    reticulate::source_python(file) # if you're missing a package, use reticulate::py_install("package_name")
  }
}

# check if all functions have been sourced
ls()
```

### {.tabset .tabset-fade .tabset-pills}
Let's perform exploratory data analysis on our dataset!

```{r}
#| include: FALSE
# this chunk is to create the plot_boundary function and import img —————

plot_boundary <- function(subzone) {
  
  border <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = T) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  return(border)
}
  
img <- "../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif"
r <- rast(img) |>
  project("EPSG:4326") 
```

#### Changi

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot changi —————

subzone <- "CHANGI"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_changi_long`:

```{r, results='hide'}
# this chunk is to pre-process changi's landsat7 & landsat8 data —————

subzone <- "CHANGI"

# landsat 7
# extract, impute and combine data from images
df_changi7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
rownames(df_changi7) <- NULL

# landsat 8
df_changi8 <- combine_img(subzone, landsat_no = 8)
rownames(df_changi8) <- NULL

merged_changi <- merge_ldst(df_changi7, df_changi8)

# aggregate temporally and obtain final data
final_changi_long <- st_aggregate(merged_changi, subzone)[[1]]
final_changi_wide <- st_aggregate(merged_changi, subzone)[[2]]
```

By using `final_changi_long`, we can obtain the following EDA results:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

eda(final_changi_long)

gif_changi <- py$create_lst_animation(final_changi_long, subzone = "CHANGI", fps = 2)
include_graphics(gif_changi)
```

To check how many NA values are present in each image, we can compute some summary statistics:

```{r, results = 'hide'}
#| include: FALSE
#| echo: TRUE
# this chunk is to write the functions for eda of missing values —————
subzones <- c("CHANGI", "JURONG EAST", "JURONG WEST")

landsat7_dir <- "../Data/Landsat/GEE_landsat7"
landsat8_dir <- "../Data/Landsat/GEE_landsat8"

landsat7_images <- list.files(path = landsat7_dir, pattern = "\\.tif$", full.names = TRUE)
landsat8_images <- list.files(path = landsat8_dir, pattern = "\\.tif$", full.names = TRUE)

all_results <- data.frame()

# processing landsat 7 images
for (img in landsat7_images) {
  result <- countNA(img, subzones, landsat_no = 7)
  result$landsat_number <- "7"
  all_results <- rbind(all_results, result)
}

# processing landsat 8 images
for (img in landsat8_images) {
  result <- countNA(img, subzones, landsat_no = 8)
  result$landsat_number <- "8"
  all_results <- rbind(all_results, result)
}

# format dates for better visualisation
all_results$formatted_date <- format(all_results$date, "%b %d, %Y")

# compute probability distribution
na_prob_distribution <- na_probdist(all_results)
```

```{r}
write.csv(all_results |>
            filter(subzone == "CHANGI"), 
          "../Data/Misc/CHANGI_NA_Results.csv", row.names = FALSE)

# summary statistics for missing values for Changi
changi_summary <- create_subzone_summary(all_results, "CHANGI")
cat("\nSummary Statistics for CHANGI:\n")
print(changi_summary)

# probability distribution
changi_prob_plot <- plot_na_probdist(na_prob_distribution, "CHANGI")
print(changi_prob_plot)
```

To better understand the distribution of the NA values, we plot the following:

```{r, results='hide'}
#| include: FALSE
# this chunk is to write the functions for eda of missing values —————

# function to create RDS images to create gif for NA values
distributionNA <- function(img, subzone) {
  date <- as.Date(str_extract(img, "\\d{4}-\\d{2}-\\d{2}"), format="%Y-%m-%d")
  
  # import subzone boundary 
  boundary <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = T) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  # import image
  r <- rast(img) |>
    project("EPSG:4326") 
  
  # align to a common grid (using LST_Singapore_2013-04-24.tif as template)
  template <- rast(paste0("../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif")) |> 
    project("EPSG:4326")
  r_aligned <- resample(r, template)
  
  # save the min. temp of region
  min_temp <- global(terra::mask(r_aligned, vect(boundary)), fun = "min", na.rm = TRUE)[1,1] 
  min_temp <- ifelse(min_temp <= 18 | is.na(min_temp), 19, min_temp)
  
  # set NA values in image to be 0
  r_aligned[is.na(r_aligned)] <- 0
  
  # mask & crop to subzone region
  temp_boundary <- terra::mask(r_aligned, vect(boundary))
  ext <- ext(boundary)
  temp_boundary <- terra::crop(temp_boundary, ext) 
  
  # aggregate spatially to reduce the number of points
  aggregated_temp_boundary <- temp_boundary |>
    aggregate(fact = 10/3, fun = "mean", na.rm = T) 
  
  # temperature lower than min_temp -> set to 0 i.e. take it as missing data
  aggregated_temp_boundary <- ifel(aggregated_temp_boundary$LST < min_temp, 0, aggregated_temp_boundary$LST)
  
  # extract as dataframe & set 0 values to be NA
  df_boundary <- terra::as.data.frame(aggregated_temp_boundary, xy = TRUE, na.rm = TRUE) |>
    mutate(LST = ifelse(LST == 0, NA, LST),
           date = date)
  
  saveRDS(df_boundary, file = paste0("../Data/Misc/SavedRDS/", gsub("\\.tif$", "", basename(img)), ".RDS"))
  print(paste0("Successfully extracted: ", basename(img)))
}

combine_distributionNA <- function(subzone, landsat_no) {
  # list all pre-existing RDS files in the target directory
  existing_rds <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
  
  # remove pre-existing RDS files if any exist
  if(length(existing_rds) > 0) {
    file.remove(existing_rds)
  }
  
  # retrieve all tif files
  ls_img <- list.files(paste0("../Data/Landsat/GEE_landsat", landsat_no), pattern = "\\.tif$", full.names = TRUE)
  
  # apply distributionNA function on all images
  lapply(ls_img, function(img) distributionNA(img, subzone))
  
  # retrieve all imputed data files
  rds_img <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
  rds_df_combined <- bind_rows(lapply(rds_img, readRDS))
  
  print(paste0("Data is combined! Ready for merging and plotting!"))
  return(rds_df_combined)
}

# function to count NA values by region
count_NA_region <- function(img, subzone) {
  
  date <- as.Date(str_extract(img, "\\d{4}-\\d{2}-\\d{2}"), format="%Y-%m-%d")
  
  boundary <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = TRUE) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  centroid <- st_centroid(boundary)
  x_mid <- st_coordinates(centroid)[1, 1]
  y_mid <- st_coordinates(centroid)[1, 2]
  
  r <- rast(img) |>
    project("EPSG:4326")
  
  template <- rast("../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif") |>
    project("EPSG:4326")
  r_aligned <- resample(r, template)
  
  min_temp <- global(terra::mask(r_aligned, vect(boundary)), fun = "min", na.rm = TRUE)[1, 1]
  min_temp <- ifelse(min_temp <= 18 | is.na(min_temp), 19, min_temp)
  r_aligned[is.na(r_aligned)] <- 0
  
  temp_boundary <- terra::mask(r_aligned, vect(boundary))
  temp_boundary <- terra::crop(temp_boundary, ext(boundary))
  
  aggregated_temp_boundary <- temp_boundary |>
    aggregate(fact = 10/3, fun = "mean", na.rm = TRUE)
  
  aggregated_temp_boundary <- ifel(aggregated_temp_boundary$LST < min_temp, 0, aggregated_temp_boundary$LST)
  
  df_boundary <- terra::as.data.frame(aggregated_temp_boundary, xy = TRUE, na.rm = TRUE) |>
    mutate(LST = ifelse(LST == 0, NA, LST),
           date = date)
  
  na_regions <- df_boundary |>
    filter(is.na(LST)) 
  
  if (nrow(na_regions) == 0) {
    
    return()  

  } else {
    
    na_regions <- na_regions |>
    mutate(region = case_when(
      x < x_mid & abs(y - y_mid) <= (max(y) - min(y)) * 0.25 ~ "West",
      x > x_mid & abs(y - y_mid) <= (max(y) - min(y)) * 0.25 ~ "East",
      y > y_mid & abs(x - x_mid) <= (max(x) - min(x)) * 0.25 ~ "North",
      y < y_mid & abs(x - x_mid) <= (max(x) - min(x)) * 0.25 ~ "South",
      TRUE ~ "Central"
    ))
  
  na_counts <- na_regions |>
    group_by(region) |>
    summarise(na_count = n()) |>
    arrange(desc(na_count))
  
  max_region <- na_counts$region[which.max(na_counts$na_count)]
  
  results <- data.frame(subzone, date, image_name = basename(img), max_region) |> cbind(na_counts)
  saveRDS(results, file = paste0("../Data/Misc/SavedRDS/", gsub("\\.tif$", "", basename(img)), ".RDS"))
  
  cat(sprintf("The maximum number of NA values is in the %s region.\n", max_region))
  }
}

ls_img <- c(list.files(paste0("../Data/Landsat/GEE_landsat", 7), pattern = "\\.tif$", full.names = TRUE), 
            list.files(paste0("../Data/Landsat/GEE_landsat", 8), pattern = "\\.tif$", full.names = TRUE))
```

```{r, results='hide'}

changi_NA7 <- combine_distributionNA("CHANGI", 7)
changi_NA8 <- combine_distributionNA("CHANGI", 8)

changi_NA <- bind_rows(changi_NA7, changi_NA8) |>
    group_by(x, y, date) |>
    summarise(LST = mean(LST, na.rm = TRUE), .groups = "drop") |> 
  mutate(missing = ifelse(is.na(LST), 1, 0))

existing_rds <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
if(length(existing_rds) > 0) { 
  file.remove(existing_rds) 
  }

lapply(ls_img, function(img) count_NA_region(img, "CHANGI"))

rds_img <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
rds_changi_combined <- bind_rows(lapply(rds_img, readRDS))
```

```{r}
#| fig.width: 10
  
gif_changi_NA <- py$create_na_animation(changi_NA, subzone = "CHANGI", value_df = rds_changi_combined, fps = 2)
include_graphics(gif_changi_NA)
```

#### Jurong East

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot jurong east —————

subzone <- "JURONG EAST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_je_long`:

```{r, results='hide'}
# this chunk is to pre-process jurong east's landsat7 & landsat8 data —————

subzone <- "JURONG EAST"

# landsat 7
# extract, impute and combine data from images
df_je7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
rownames(df_je7) <- NULL

# landsat 8
df_je8 <- combine_img(subzone, landsat_no = 8)
rownames(df_je8) <- NULL

merged_je <- merge_ldst(df_je7, df_je8)

# aggregate temporally and obtain final data
final_je_long <- st_aggregate(merged_je, subzone)[[1]]
final_je_wide <- st_aggregate(merged_je, subzone)[[2]]
```

By using `final_je_long`, we can obtain the following EDA results:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

eda(final_je_long)

gif_je <- py$create_lst_animation(final_je_long, subzone = "JURONG EAST", fps = 2)
include_graphics(gif_je)
```

To check how many NA values are present in each image, we can compute some summary statistics:

```{r}
write.csv(all_results |>
            filter(subzone == "JURONG EAST"), 
          "../Data/Misc/JURONG EAST_NA_Results.csv", row.names = FALSE)

# summary statistics for missing values for Jurong East
je_summary <- create_subzone_summary(all_results, "JURONG EAST")
cat("\nSummary Statistics for JURONG EAST:\n")
print(je_summary)

# probability distribution
je_prob_plot <- plot_na_probdist(na_prob_distribution, "JURONG EAST")
print(je_prob_plot)
```

To better understand the distribution of the NA values, we plot the following:

```{r, results='hide'}
je_NA7 <- combine_distributionNA("JURONG EAST", 7)
je_NA8 <- combine_distributionNA("JURONG EAST", 8)

je_NA <- bind_rows(je_NA7, je_NA8) |>
    group_by(x, y, date) |>
    summarise(LST = mean(LST, na.rm = TRUE), .groups = "drop") |> 
  mutate(missing = ifelse(is.na(LST), 1, 0))

existing_rds <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
if(length(existing_rds) > 0) { 
  file.remove(existing_rds) 
  }

lapply(ls_img, function(img) count_NA(img, "JURONG EAST"))

rds_img <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
rds_je_combined <- bind_rows(lapply(rds_img, readRDS))
```

```{r}
#| fig.width: 10

gif_je_NA <- py$create_na_animation(je_NA, subzone = "JURONG EAST", value_df = rds_je_combined, fps = 2)
include_graphics(gif_je_NA)
```

#### Jurong West

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot jurong west —————

subzone <- "JURONG WEST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_jw_long`:

```{r, results='hide'}
# this chunk is to pre-process jurong west's landsat7 & landsat8 data —————

subzone <- "JURONG WEST"

# landsat 7
# extract, impute and combine data from images
df_jw7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
rownames(df_jw7) <- NULL

# landsat 8
df_jw8 <- combine_img(subzone, landsat_no = 8)
rownames(df_jw8) <- NULL

merged_jw <- merge_ldst(df_jw7, df_jw8)

# aggregate temporally and obtain final data
final_jw_long <- st_aggregate(merged_jw, subzone)[[1]]
final_jw_wide <- st_aggregate(merged_jw, subzone)[[2]]
```

By using `final_jw_long`, we can see how the LST values changes over time:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

eda(final_jw_long)

gif_jw <- py$create_lst_animation(final_jw_long, subzone = "JURONG WEST", fps = 2)
include_graphics(gif_jw)
```

To check how many NA values are present in each image, we can compute some summary statistics:

```{r}
write.csv(all_results |>
            filter(subzone == "JURONG WEST"), 
          "../Data/Misc/JURONG EAST_NA_Results.csv", row.names = FALSE)

# summary statistics for missing values for Jurong West
jw_summary <- create_subzone_summary(all_results, "JURONG WEST")
cat("\nSummary Statistics for JURONG WEST:\n")
print(jw_summary)

# probability distribution
jw_prob_plot <- plot_na_probdist(na_prob_distribution, "JURONG WEST")
print(jw_prob_plot)
```

To better understand the distribution of the NA values, we plot the following:

```{r, results='hide'}
jw_NA7 <- combine_distributionNA("JURONG WEST", 7)
jw_NA8 <- combine_distributionNA("JURONG WEST", 8)

jw_NA <- bind_rows(jw_NA7, jw_NA8) |>
    group_by(x, y, date) |>
    summarise(LST = mean(LST, na.rm = TRUE), .groups = "drop") |> 
  mutate(missing = ifelse(is.na(LST), 1, 0))

existing_rds <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
if(length(existing_rds) > 0) { 
  file.remove(existing_rds) 
  }

lapply(ls_img, function(img) count_NA(img, "JURONG WEST"))

rds_img <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
rds_jw_combined <- bind_rows(lapply(rds_img, readRDS))
```

```{r}
#| fig.width: 10

gif_jw_NA <- py$create_na_animation(jw_NA, subzone = "JURONG WEST", value_df = rds_jw_combined, fps = 2)
include_graphics(gif_jw_NA)
```
