---
title: "ST5188 PG01 Exploratory Data Analysis"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r}
#| include: FALSE
# this chunk is to source all our functions —————
# Clear R environment safely

# Load necessary spatial packages
library(terra)
library(sf)
library(dplyr)
library(here)
library(data.table)
library(pbapply)

# clear working directory
rm(list = ls())
scripts = list.files("../Code/Preprocessing", full.names = TRUE)

reticulate::py_config()
invisible(reticulate::py_install("matplotlib"))
invisible(reticulate::py_install("imageio"))
invisible(reticulate::py_install("geopandas"))
# source libraries and functions under `Code` files
for (file in scripts) {
  print(paste("Sourcing:", file))
  if (grepl("\\.R$", file)) {
    source(file)
  } else {
    print(paste("Python sourcing:", file))
    reticulate::source_python(file)
  }
}

# check if all functions have been sourced
ls()

```

### {.tabset .tabset-fade .tabset-pills}
Let's perform exploratory data analysis on our dataset!

```{r Singapore shapefile}
#| include: FALSE
# this chunk is to create the plot_boundary function and import img —————

# Function to plot boundary for any subzone
plot_boundary <- function(subzone) {
  
  # Load subzone shapefile
   border <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", 
                     quiet = TRUE) |>
    select(PLN_AREA_N, geometry) |> 
    filter(PLN_AREA_N == subzone) |> 
    st_union() |> 
    st_transform("EPSG:4326") |> 
    st_sf()
  
  return(border)
}

# Example Image Import: Single Landsat 8 image
img <- "../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif"
r <- rast(img) |>
  project("EPSG:4326") 

```

#### Changi

We will be focusing on the LST values in the blue border below:

```{r Changi plot}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot Changi —————
subzone <- "CHANGI"

# Plot Base Raster (Landsat Image)
plot(r)

# Overlay Boundary with Transparency
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))

```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_changi_long`:

```{r Preprocessing}
#| results='hide'
# this chunk is to pre-process Changi's landsat7 & landsat8 data and SAVE the result —————

subzone <- "CHANGI"

# Define file paths to save the processed data RELATIVE to the Rmd file
output_file_long <- file.path("../Data/Final/final_changi_long.csv")
output_file_wide <- file.path("../Data/Final/final_changi_wide.csv")

# Check if the saved files already exist
if (!file.exists(output_file_long) || !file.exists(output_file_wide)) {
  message("Pre-processed data for CHANGI not found. Running preprocessing...")
  # landsat 7
  df_changi7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
  rownames(df_changi7) <- NULL

  # landsat 8
  df_changi8 <- combine_img(subzone, landsat_no = 8)
  rownames(df_changi8) <- NULL

  merged_changi <- merge_ldst(df_changi7, df_changi8)

  # aggregate temporally and obtain final data
  final_changi_long <- st_aggregate(merged_changi, subzone)[[1]]
  final_changi_wide <- st_aggregate(merged_changi, subzone)[[2]]

  # Save the results to CSV files RELATIVE to the Rmd file
  write.csv(final_changi_long, output_file_long, row.names = FALSE)
  write.csv(final_changi_wide, output_file_wide, row.names = FALSE)
  message("Pre-processed data for CHANGI saved to CSV files.")
} else {
  message("Pre-processed data for CHANGI already exists. Skipping preprocessing.")
  # The data will be loaded in a subsequent chunk
}
```

```{r Loading preprocessed data}
# this chunk is to LOAD the pre-processed changi data —————

# Define file path RELATIVE to the Rmd file (same as saving path)
output_file_long <- file.path("../Data/Final/final_changi_long.csv")

if (file.exists(output_file_long)) {
  final_changi_long <- read.csv(output_file_long)
  # Ensure 'date' column is treated as Date if it exists
  if ("date" %in% colnames(final_changi_long)) {
    final_changi_long$date <- as.Date(final_changi_long$date)
  }
  print("Pre-processed CHANGI data loaded successfully.")
} else {
  print("Pre-processed CHANGI data file not found. Ensure the preprocessing chunk has been run at least once to save the data.")
}
```

By using `final_changi_long`, we can obtain the following EDA results:

```{r eda:Changi}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis and save the animation if not generated else it skips the generation steps retrieves th existing from previous run—————

# Ensure the 'eda' function is defined and sourced
eda(final_changi_long)

# Define the output GIF file path
output_gif_changi <- ("../Data/Misc/gif/LST_animation_CHANGI.gif")

# Check if the GIF file already exists
if (!file.exists(output_gif_changi)) {
  message("Generating LST animation for CHANGI...")
  gif_changi <- py$create_lst_animation(final_changi_long, subzone = "CHANGI", fps = 2, output_path = output_gif_changi)
  message(paste("LST animation saved to:", output_gif_changi))
} else {
  message("LST animation for CHANGI already exists. Skipping generation.")
  gif_changi <- output_gif_changi # Use the existing file path
}

include_graphics(gif_changi)
```

To check how many NA values are present in each image, we can compute some summary statistics:




```{r missing values analysis:Changi, results='hide', include=FALSE}
#| echo: TRUE
#| message: TRUE
#| warning: TRUE
#| cache: FALSE 

### EDA of Missing Values - Optimized with Caching ###

output_dir <- "../Data/Misc"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

# Define output files
output_all_results <- file.path(output_dir, "all_results.rds")  # Using RDS for better preservation
output_na_prob_dist <- file.path(output_dir, "na_prob_dist.rds")

# Check for cached results
if (file.exists(output_all_results) && file.exists(output_na_prob_dist)) {
  message("Loading cached results...")
  all_results <- readRDS(output_all_results)
  na_prob_distribution <- readRDS(output_na_prob_dist)
} else {
  # Define subzones
  subzones <- c("CHANGI", "JURONG EAST", "JURONG WEST")
  
  # Data Loading with Validation
  landsat7_dir <- "../Data/Landsat/GEE_landsat7"
  landsat8_dir <- "../Data/Landsat/GEE_landsat8"
  
  stopifnot("Landsat7 directory missing" = dir.exists(landsat7_dir),
            "Landsat8 directory missing" = dir.exists(landsat8_dir))
  
  message("Listing image files...")
  landsat7_images <- list.files(landsat7_dir, "\\.tif$", full.names = TRUE)
  landsat8_images <- list.files(landsat8_dir, "\\.tif$", full.names = TRUE)
  
  # Process images with progress bar
  process_images <- function(images, landsat_no) {
    if (length(images) == 0) return(data.table())
    rbindlist(pblapply(images, function(img) {
      result <- as.data.table(countNA(img, subzones, landsat_no))
      result[, landsat_number := landsat_no]
      result
    }))
  }
  
  message("Processing images...")
  ls7_results <- process_images(landsat7_images, 7)
  ls8_results <- process_images(landsat8_images, 8)
  
  # Combine Results
  all_results <- rbindlist(list(ls7_results, ls8_results), fill = TRUE)
  
  # Post-processing
  if (nrow(all_results) > 0) {
    all_results[, formatted_date := format(date, "%b %d, %Y")]
    na_prob_distribution <- na_probdist(all_results)
    
    # Save results
    saveRDS(all_results, output_all_results)
    saveRDS(na_prob_distribution, output_na_prob_dist)
    message("Results saved to cache.")
  } else {
    stop("No results generated - check input data")
  }
}

# Show results
message("\nCurrent results summary:")
print(all_results[, .(mean_na = mean(na_count/total_pixels)), by = landsat_number])
```


```{r summary statistics for missing values:Changi}
write.csv(all_results |>
            filter(subzone == "CHANGI"), 
          "../Data/Misc/CHANGI_NA_Results.csv", row.names = FALSE)

# summary statistics for missing values for Changi
changi_summary <- create_subzone_summary(all_results, "CHANGI")
cat("\nSummary Statistics for CHANGI:\n")
print(changi_summary)

# probability distribution
changi_prob_plot <- plot_na_probdist(na_prob_distribution, "CHANGI")
print(changi_prob_plot)
```

To better understand the distribution of the NA values, we plot the following:

```{r distribution of NA values}
#| include: FALSE
# this chunk is to write the functions for eda of missing values —————

# function to create RDS images to create gif for NA values
distributionNA <- function(img, subzone) {
  date <- as.Date(str_extract(img, "\\d{4}-\\d{2}-\\d{2}"), format="%Y-%m-%d")
  
  # import subzone boundary 
  boundary <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = T) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  # import image
  r <- rast(img) |>
    project("EPSG:4326") 
  
  # align to a common grid (using LST_Singapore_2013-04-24.tif as template)
  template <- rast(paste0("../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif")) |> 
    project("EPSG:4326")
  r_aligned <- resample(r, template)
  
  # save the min. temp of region
  min_temp <- global(terra::mask(r_aligned, vect(boundary)), fun = "min", na.rm = TRUE)[1,1] 
  min_temp <- ifelse(min_temp <= 18 | is.na(min_temp), 19, min_temp) # to account for cases where temperatures are negative or extremely low
  
  # set NA values in image to be 0
  r_aligned[is.na(r_aligned)] <- 0
  
  # mask & crop to subzone region
  temp_boundary <- terra::mask(r_aligned, vect(boundary))
  ext <- ext(boundary)
  temp_boundary <- terra::crop(temp_boundary, ext) 
  
  # aggregate spatially to reduce the number of points
  aggregated_temp_boundary <- temp_boundary |>
    aggregate(fact = 10/3, fun = "mean", na.rm = T) 
  
  # temperature lower than min_temp -> set to 0 i.e. take it as missing data
  aggregated_temp_boundary <- ifel(aggregated_temp_boundary$LST < min_temp, 0, aggregated_temp_boundary$LST)
  
  # extract as dataframe & set 0 values to be NA
  df_boundary <- terra::as.data.frame(aggregated_temp_boundary, xy = TRUE, na.rm = TRUE) |>
    mutate(LST = ifelse(LST == 0, NA, LST),
           date = date)
  
  saveRDS(df_boundary, file = paste0("../Data/Misc/SavedRDS/", gsub("\\.tif$", "", basename(img)), ".RDS"))
  print(paste0("Successfully extracted: ", basename(img)))
}

combine_distributionNA <- function(subzone, landsat_no) {
  # list all pre-existing RDS files in the target directory
  existing_rds <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
  
  # remove pre-existing RDS files if any exist
  if(length(existing_rds) > 0) {
    file.remove(existing_rds)
  }
  
  # retrieve all tif files
  ls_img <- list.files(paste0("../Data/Landsat/GEE_landsat", landsat_no), pattern = "\\.tif$", full.names = TRUE)
  
  # apply distributionNA function on all images
  lapply(ls_img, function(img) distributionNA(img, subzone))
  
  # retrieve all imputed data files
  rds_img <- list.files("../Data/Misc/SavedRDS", pattern = "\\.RDS$", full.names = TRUE)
  rds_df_combined <- bind_rows(lapply(rds_img, readRDS))
  
  print(paste0("Data is combined! Ready for merging and plotting!"))
  return(rds_df_combined)
}

# function to count NA values by region
count_NA_region <- function(img, subzone) {
  date <- as.Date(str_extract(img, "\\d{4}-\\d{2}-\\d{2}"), format="%Y-%m-%d")
  
  boundary <- st_read("../Data/Misc/Subzone/MP14_SUBZONE_NO_SEA_PL.shp", quiet = TRUE) |>
    select(PLN_AREA_N, geometry) |>
    filter(PLN_AREA_N == subzone) |>
    st_union() |>
    st_transform("EPSG:4326") |>
    st_sf()
  
  centroid <- st_centroid(boundary)
  x_mid <- st_coordinates(centroid)[1, 1]
  y_mid <- st_coordinates(centroid)[1, 2]
  
  r <- rast(img) |>
    project("EPSG:4326")
  
  template <- rast("../Data/Landsat/GEE_landsat8/LST_Singapore_2013-04-24.tif") |>
    project("EPSG:4326")
  r_aligned <- resample(r, template)
  
  min_temp <- global(terra::mask(r_aligned, vect(boundary)), fun = "min", na.rm = TRUE)[1, 1]
  min_temp <- ifelse(min_temp <= 18 | is.na(min_temp), 19, min_temp)
  r_aligned[is.na(r_aligned)] <- 0
  
  temp_boundary <- terra::mask(r_aligned, vect(boundary))
  temp_boundary <- terra::crop(temp_boundary, ext(boundary))
  
  aggregated_temp_boundary <- temp_boundary |>
    aggregate(fact = 10/3, fun = "mean", na.rm = TRUE)
  
  aggregated_temp_boundary <- ifel(aggregated_temp_boundary$LST < min_temp, 0, aggregated_temp_boundary$LST)
  
  df_boundary <- terra::as.data.frame(aggregated_temp_boundary, xy = TRUE, na.rm = TRUE) |>
    mutate(LST = ifelse(LST == 0, NA, LST),
           date = date)
  
  # Skip processing if there are no missing values
  if (sum(is.na(df_boundary$LST)) == 0) {
    cat(sprintf("No missing values found in %s for %s. Skipping...\n", basename(img), subzone))
    return(NULL)  # Skip this image if no NA values
  }
  
  na_regions <- df_boundary |>
    filter(is.na(LST)) |>
    mutate(region = case_when(
      x < x_mid & abs(y - y_mid) <= (max(df_boundary$y, na.rm = TRUE) - min(df_boundary$y, na.rm = TRUE)) * 0.25 ~ "West",
      x > x_mid & abs(y - y_mid) <= (max(df_boundary$y, na.rm = TRUE) - min(df_boundary$y, na.rm = TRUE)) * 0.25 ~ "East",
      y > y_mid & abs(x - x_mid) <= (max(df_boundary$x, na.rm = TRUE) - min(df_boundary$x, na.rm = TRUE)) * 0.25 ~ "North",
      y < y_mid & abs(x - x_mid) <= (max(df_boundary$x, na.rm = TRUE) - min(df_boundary$x, na.rm = TRUE)) * 0.25 ~ "South",
      TRUE ~ "Central"
    ))
  
  na_counts <- na_regions |>
    group_by(region) |>
    summarise(na_count = n()) |>
    arrange(desc(na_count))
  
  # Handle case where na_counts is empty
  if (nrow(na_counts) == 0) {
    max_region <- "None"  # No missing data
    na_counts <- data.frame(region = "None", na_count = 0)
  } else {
    max_region <- na_counts$region[which.max(na_counts$na_count)]
  }
  
  results <- data.frame(subzone, date, image_name = basename(img), max_region, na_counts)
  saveRDS(results, file = paste0("../Data/Misc/SavedRDS/", gsub("\\.tif$", "", basename(img)), ".RDS"))
  
  cat(sprintf("The maximum number of NA values is in the %s region.\n", max_region))
}

ls_img <- c(list.files(paste0("../Data/Landsat/GEE_landsat", 7), pattern = "\\.tif$", full.names = TRUE), 
            list.files(paste0("../Data/Landsat/GEE_landsat", 8), pattern = "\\.tif$", full.names = TRUE))
```


```{r distribution of NA values :Changi, warning=FALSE,   echo=TRUE}

subzone <- "CHANGI"
cache_dir <- "../Data/Misc/SavedRDS"
dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)

# Define cache files
combined_cache <- file.path(cache_dir, "changi_combined_data.RDS")
na_counts_cache <- file.path(cache_dir, "changi_na_counts.RDS")

# Check if cached results exist
if (file.exists(combined_cache) && file.exists(na_counts_cache)) {
  message("Using the cached results.")
  changi_NA <- readRDS(combined_cache)
  rds_changi_combined <- readRDS(na_counts_cache)
} else {
  message("As no cached results found, processing data")
  
  # Process combined distributions
  changi_NA7 <- combine_distributionNA(subzone, 7)
  changi_NA8 <- combine_distributionNA(subzone, 8)
  
  changi_NA <- bind_rows(changi_NA7, changi_NA8) |>
    group_by(x, y, date) |>
    summarise(LST = mean(LST, na.rm = TRUE), .groups = "drop") |> 
    mutate(missing = ifelse(is.na(LST), 1, 0))
  
  # Process NA counts
  temp_rds_dir <- file.path(cache_dir, "temp")
  dir.create(temp_rds_dir, showWarnings = FALSE)
  
  # Clear only temporary files, not the main cache
  temp_files <- list.files(temp_rds_dir, full.names = TRUE)
  if (length(temp_files) > 0) file.remove(temp_files)
  
  # Get image files
  ls_img <- c(
    list.files("../Data/Landsat/GEE_landsat7", "\\.tif$", full.names = TRUE),
    list.files("../Data/Landsat/GEE_landsat8", "\\.tif$", full.names = TRUE)
  )
  
  # Process with progress
  pbapply::pblapply(ls_img, function(img) {
    count_NA_region(img, subzone)
  })
  
  # Combine results
  rds_img <- list.files(temp_rds_dir, full.names = TRUE)
  rds_changi_combined <- bind_rows(lapply(rds_img, readRDS))
  
  # Save to cache
  saveRDS(changi_NA, combined_cache)
  saveRDS(rds_changi_combined, na_counts_cache)
  
  # Clean up temp files
  unlink(temp_rds_dir, recursive = TRUE)
}

# Verify results
stopifnot("No results generated" = exists("changi_NA") && exists("rds_changi_combined"))
```


```{r Plotting NA values: Changi}
#| fig.width: 10

# Define the output GIF file path for CHANGI NA animation
output_gif_changi_NA <- "../Data/Misc/gif/NA_animation_CHANGI.gif"



# Check if the GIF file already exists
if (!file.exists(output_gif_changi_NA)) {
  gif_changi_NA <- py$create_na_animation(changi_NA, subzone = "CHANGI", value_df = rds_changi_combined, fps = 2, output_path = output_gif_changi_NA)
  message(paste("NA animation for CHANGI saved to:", output_gif_changi_NA))
} else {
  message("NA animation for CHANGI already exists. Skipping generation.")
  gif_changi_NA <- output_gif_changi_NA
  
  
}

include_graphics(gif_changi_NA)
```

#### Jurong East

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot jurong east —————

subzone <- "JURONG EAST"

# Plot Base Raster (Landsat Image)
plot(r)

# Overlay Boundary with Transparency
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_je_long`:

```{r}
#| results='hide'
# this chunk is to pre-process Jurong East's landsat7 & landsat8 data and SAVE the result —————

subzone <- "JURONG EAST"

# Define file paths to save the processed data RELATIVE to the Rmd file
output_file_long <- file.path("../Data/Final/final_jurong east_long.csv")
output_file_wide <- file.path("../Data/Final/final_jurong east_wide.csv")

# Check if the saved files already exist
if (!file.exists(output_file_long) || !file.exists(output_file_wide)) {
  message("Pre-processed data for JURONG EAST not found. Running preprocessing...")
  # landsat 7
  df_je7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
  rownames(df_je7) <- NULL

  # landsat 8
  df_je8 <- combine_img(subzone, landsat_no = 8)
  rownames(df_je8) <- NULL

  merged_je <- merge_ldst(df_je7, df_je8)

  # aggregate temporally and obtain final data
  final_je_long <- st_aggregate(merged_je, subzone)[[1]]
  final_je_wide <- st_aggregate(merged_je, subzone)[[2]]

  # Save the results to CSV files RELATIVE to the Rmd file
  write.csv(final_je_long, output_file_long, row.names = FALSE)
  write.csv(final_je_wide, output_file_wide, row.names = FALSE)
  message("Pre-processed data for JURONG EAST saved to CSV files.")
} else {
  message("Pre-processed data for JURONG EAST already exists. Skipping preprocessing.")
  # The data will be loaded in a subsequent chunk
}
```

```{r}
# this chunk is to LOAD the pre-processed Jurong East data —————

output_file_long <- file.path("../Data/Final/final_jurong east_long.csv")

if (file.exists(output_file_long)) {
  final_je_long <- read.csv(output_file_long)
  # Ensure 'date' column is treated as Date if it exists
  if ("date" %in% colnames(final_je_long)) {
    final_je_long$date <- as.Date(final_je_long$date)
  }
  print("Pre-processed JURONG EAST data loaded successfully.")
} else {
  print("Pre-processed JURONG EAST data file not found. Ensure the preprocessing chunk has been run at least once to save the data.")
}
```

By using `final_je_long`, we can obtain the following EDA results:



```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis and save the animation if not generated else it skips the generation steps retrieves th existing from previous run—————

# Ensure the 'eda' function is defined and sourced
eda(final_je_long)

# Define the output GIF file path
output_gif_je <- ("../Data/Misc/gif/LST_animation_JURONG EAST.gif")

# Check if the GIF file already exists
if (!file.exists(output_gif_je)) {
  message("Generating LST animation for JURONG EAST...")
  gif_je <- py$create_lst_animation(final_je_long, subzone = "JURONG EAST", fps = 2, output_path = output_gif_je)
  message(paste("LST animation saved to:", output_gif_je))
} else {
  message("LST animation for JURONG EAST already exists. Skipping generation.")
  gif_je <- output_gif_je # Use the existing file path
}

include_graphics(gif_je)
```


To check how many NA values are present in each image, we can compute some summary statistics:

```{r}
write.csv(all_results |>
            filter(subzone == "JURONG EAST"), 
          "../Data/Misc/JURONG EAST_NA_Results.csv", row.names = FALSE)

# summary statistics for missing values for Jurong East
je_summary <- create_subzone_summary(all_results, "JURONG EAST")
cat("\nSummary Statistics for JURONG EAST:\n")
print(je_summary)

# probability distribution
je_prob_plot <- plot_na_probdist(na_prob_distribution, "JURONG EAST")
print(je_prob_plot)
```

To better understand the distribution of the NA values, we plot the following:

```{r}
subzone <- "JURONG EAST"
cache_dir <- "../Data/Misc/SavedRDS"
dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)

# Define cache files
combined_cache <- file.path(cache_dir, "je_combined_data.RDS")
na_counts_cache <- file.path(cache_dir, "je_na_counts.RDS")

# Check if cached results exist
if (file.exists(combined_cache) && file.exists(na_counts_cache)) {
  message("Using the cached results.")
  je_NA <- readRDS(combined_cache)
  rds_je_combined <- readRDS(na_counts_cache)
} else {
  message("As no cached results found, processing data")
  
  # Process combined distributions
  je_NA7 <- combine_distributionNA(subzone, 7)
  je_NA8 <- combine_distributionNA(subzone, 8)
  
  je_NA <- bind_rows(je_NA7, je_NA8) |>
    group_by(x, y, date) |>
    summarise(LST = mean(LST, na.rm = TRUE), .groups = "drop") |> 
    mutate(missing = ifelse(is.na(LST), 1, 0))
  
  # Process NA counts
  temp_rds_dir <- file.path(cache_dir, "temp")
  dir.create(temp_rds_dir, showWarnings = FALSE)
  
  # Clear only temporary files, not the main cache
  temp_files <- list.files(temp_rds_dir, full.names = TRUE)
  if (length(temp_files) > 0) file.remove(temp_files)
  
  # Get image files
  ls_img <- c(
    list.files("../Data/Landsat/GEE_landsat7", "\\.tif$", full.names = TRUE),
    list.files("../Data/Landsat/GEE_landsat8", "\\.tif$", full.names = TRUE)
  )
  
  # Process with progress
  pbapply::pblapply(ls_img, function(img) {
    count_NA_region(img, subzone)
  })
  
  # Combine results
  rds_img <- list.files(temp_rds_dir, full.names = TRUE)
  rds_je_combined <- bind_rows(lapply(rds_img, readRDS))
  
  # Save to cache
  saveRDS(je_NA, combined_cache)
  saveRDS(rds_je_combined, na_counts_cache)
  
  # Clean up temp files
  unlink(temp_rds_dir, recursive = TRUE)
}

# Verify results
stopifnot("No results generated" = exists("je_NA") && exists("rds_je_combined"))
```

```{r}
#| fig.width: 10


# Define the output GIF file path for JURONG EAST NA animation
output_gif_je_NA <- "../Data/Misc/gif/NA_animation_JURONG EAST.gif"

# To check if the GIF file already exists
if (!file.exists(output_gif_je_NA)) {
  gif_je_NA <- py$create_na_animation(je_NA, subzone = "JURONG EAST", value_df = rds_je_combined, fps = 2, output_path = output_gif_je_NA)
  message(paste("NA animation for JURONG EAST saved to:", output_gif_je_NA))
} else {
  message("NA animation for JURONG EAST already exists. Skipping generation.")
  gif_je_NA <- output_gif_je_NA
}

include_graphics(gif_je_NA)
```

#### Jurong West

We will be focusing on the LST values in the blue border below:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to plot jurong west —————

subzone <- "JURONG WEST"

plot(r)
plot(plot_boundary(subzone), add = TRUE,  col = rgb(0, 0, 1, alpha = 0.4))
```

First, we preprocessed our data using the following codes, and obtained a dataframe called `final_jw_long`:

```{r, results='hide'}
# this chunk is to pre-process jurong west's landsat7 & landsat8 data

subzone <- "JURONG WEST"

# Define file paths to save the processed data RELATIVE to the Rmd file
output_file_long <- file.path("../Data/Final/final_jurong west_long.csv")
output_file_wide <- file.path("../Data/Final/final_jurong west_wide.csv")

# Check if the saved files already exist
if (!file.exists(output_file_long) || !file.exists(output_file_wide)) {
  message("Pre-processed data for JURONG WEST not found. Running preprocessing")
  # landsat 7
  df_jw7 <- combine_img(subzone, landsat_no = 7) # takes around 3-4 mins
  rownames(df_jw7) <- NULL

  # landsat 8
  df_jw8 <- combine_img(subzone, landsat_no = 8)
  rownames(df_jw8) <- NULL

  merged_jw <- merge_ldst(df_jw7, df_jw8)

  # aggregate temporally and obtain final data
  final_jw_long <- st_aggregate(merged_jw, subzone)[[1]]
  final_jw_wide <- st_aggregate(merged_jw, subzone)[[2]]

  # Save the results to CSV files RELATIVE to the Rmd file
  write.csv(final_jw_long, output_file_long, row.names = FALSE)
  write.csv(final_jw_wide, output_file_wide, row.names = FALSE)
  message("Pre-processed data for JURONG WEST saved to CSV files.")
} else {
  message("Pre-processed data for JURONG WEST already exists. Skipping preprocessing.")
  # The data will be loaded in a subsequent chunk
}
```

```{r}
# this chunk is to LOAD the pre-processed Jurong East data —————

output_file_long <- file.path("../Data/Final/final_jurong west_long.csv")

if (file.exists(output_file_long)) {
  final_jw_long <- read.csv(output_file_long)
  # Ensure 'date' column is treated as Date if it exists
  if ("date" %in% colnames(final_jw_long)) {
    final_jw_long$date <- as.Date(final_jw_long$date)
  }
  print("Pre-processed JURONG WEST data loaded successfully.")
} else {
  print("Pre-processed JURONG WEST data file not found. Ensure the preprocessing chunk has been run at least once to save the data.")
}
```


By using `final_jw_long`, we can see how the LST values changes over time:

```{r}
#| echo: FALSE
#| fig.width: 10
# this chunk is to conduct exploratory data analysis —————

# Ensure the 'eda' function is defined and sourced
eda(final_jw_long)

# Define the output GIF file path
output_gif_jw <- ("../Data/Misc/gif/LST_animation_JURONG WEST.gif")

# Check if the GIF file already exists
if (!file.exists(output_gif_jw)) {
  message("Generating LST animation for JURONG WEST...")
  gif_jw <- py$create_lst_animation(final_jw_long, subzone = "JURONG WEST", fps = 2, output_path = output_gif_jw)
  message(paste("LST animation saved to:", output_gif_jw))
} else {
  message("LST animation for JURONG WEST already exists. Skipping generation.")
  gif_jw <- output_gif_jw # Use the existing file path
}

include_graphics(gif_jw)

```

To check how many NA values are present in each image, we can compute some summary statistics:

```{r}
write.csv(all_results |>
            filter(subzone == "JURONG WEST"), 
          "../Data/Misc/JURONG EAST_NA_Results.csv", row.names = FALSE)

# summary statistics for missing values for Jurong West
jw_summary <- create_subzone_summary(all_results, "JURONG WEST")
cat("\nSummary Statistics for JURONG WEST:\n")
print(jw_summary)

# probability distribution
jw_prob_plot <- plot_na_probdist(na_prob_distribution, "JURONG WEST")
print(jw_prob_plot)
```

To better understand the distribution of the NA values, we plot the following:

```{r, results='hide'}
subzone <- "JURONG WEST"
cache_dir <- "../Data/Misc/SavedRDS"
dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)

# Define cache files
combined_cache <- file.path(cache_dir, "jw_combined_data.RDS")
na_counts_cache <- file.path(cache_dir, "jw_na_counts.RDS")

# Check if cached results exist
if (file.exists(combined_cache) && file.exists(na_counts_cache)) {
  message("Using the cached results.")
  jw_NA <- readRDS(combined_cache)
  rds_jw_combined <- readRDS(na_counts_cache)
} else {
  message("As no cached results found, processing data")
  
  # Process combined distributions
  jw_NA7 <- combine_distributionNA(subzone, 7)
  jw_NA8 <- combine_distributionNA(subzone, 8)
  
  jw_NA <- bind_rows(jw_NA7, jw_NA8) |>
    group_by(x, y, date) |>
    summarise(LST = mean(LST, na.rm = TRUE), .groups = "drop") |> 
    mutate(missing = ifelse(is.na(LST), 1, 0))
  
  # Process NA counts
  temp_rds_dir <- file.path(cache_dir, "temp")
  dir.create(temp_rds_dir, showWarnings = FALSE)
  
  # Clear only temporary files, not the main cache
  temp_files <- list.files(temp_rds_dir, full.names = TRUE)
  if (length(temp_files) > 0) file.remove(temp_files)
  
  # Get image files
  ls_img <- c(
    list.files("../Data/Landsat/GEE_landsat7", "\\.tif$", full.names = TRUE),
    list.files("../Data/Landsat/GEE_landsat8", "\\.tif$", full.names = TRUE)
  )
  
  # Process with progress
  pbapply::pblapply(ls_img, function(img) {
    count_NA_region(img, subzone)
  })
  
  # Combine results
  rds_img <- list.files(temp_rds_dir, full.names = TRUE)
  rds_jw_combined <- bind_rows(lapply(rds_img, readRDS))
  
  # Save to cache
  saveRDS(jw_NA, combined_cache)
  saveRDS(rds_jw_combined, na_counts_cache)
  
  # Clean up temp files
  unlink(temp_rds_dir, recursive = TRUE)
}

# Verify results
stopifnot("No results generated" = exists("jw_NA") && exists("rds_jw_combined"))
```
```

```{r}
#| fig.width: 10

gif_jw_NA <- py$create_na_animation(jw_NA, subzone = "JURONG WEST", value_df = rds_jw_combined, fps = 2)
include_graphics(gif_jw_NA)
```
