---
title: "ST5188-working of spTimer"
output: html_document
---

```{r setup, include=FALSE,message=FALSE}

# clear working directory
rm(list = ls())

scripts = list.files("../Code/GP-models/spTimer", full.names = TRUE)

library(spTimer)

#memory.limit(size = 8000)  # Increase memory limit (in MB) In windows


# Load the package
library(ff)
library(dplyr)
```

Loading the data set: Changi region

```{r}
data<-as.data.frame(read.csv("/Users/varshap/Downloads/ST5188/Data/Final/TT Split/changi_train_long.csv"))
#data <- na.omit(data)  
head(data)

data <- data %>%
  mutate(Date = as.factor(Date)) %>% # Ensure Date is a factor first
  mutate(time_index = as.integer(factor(Date, levels = unique(Date))))

# Check the structure of the data
str(data)

# Check the range of time_index
range(data$time_index)  # Should return the minimum and maximum time indices

# Check the distribution of time_index
table(data$time_index)  # Should show the count of observations for each time point



# # Aggregate LST values for each (x, y) pair over time
# aggregated_data <- data %>%
#   group_by(x, y) %>%
#   summarise(
#     mean_LST = mean(Value, na.rm = TRUE),  # Mean LST
#   )
# 
# # View the aggregated data
# head(aggregated_data)
# 

# # Convert Date to a factor and create a time index
# data <- data %>%
#   mutate(
#     Date = as.factor(Date),  # Convert Date to a factor
#     time_index = as.integer(factor(Date, levels = unique(Date))))  # Create a numeric time index
# 

#Longitude:x
#Latitude:y
#time: period

```

```{r}
# Convert 'period' to a factor variable
data$Date <- as.factor(data$Date)

# Check the structure again to confirm 'period' is now a factor
str(data)

```


### using Wide format here??!
```{r}

# Load the dataset
data <- read.csv("/Users/varshap/Downloads/ST5188/Data/Final/TT Split/changi_train_wide.csv")

# Randomly select 1000 rows
set.seed(123)  # Ensures reproducibility
data <- data %>% sample_n(1000)

# Add small random noise to coordinates
set.seed(123)
data <- data %>%
  mutate(
    x = x + runif(n(), -0.1, 0.1),  # Add noise to x
    y = y + runif(n(), -0.1, 0.1)   # Add noise to y
  )

#rescaling
# data <- data %>%
#   mutate(
#     x = (x - min(x)) / (max(x) - min(x)),
#     y = (y - min(y)) / (max(y) - min(y))
#   )

# Extract coordinates
coords <- as.matrix(data[, c("x", "y")])

# Define the formula for the first time point
formula <- `Mar.Apr.2000` ~ x + y

# Define priors
priors <- spTimer::spT.priors(
  model = "GP",
  inv.var.prior = spTimer::Gamm(2, 0.1),
  beta.prior = spTimer::Norm(0, 10^4)
)

priors <- spTimer::spT.priors(
  model = "GP",
  inv.var.prior = spTimer::Gamm(2, 0.1),  # Adjust the shape and scale
  beta.prior = spTimer::Norm(0, 10^4),     # Increase variance
  phi.prior = spTimer::Gam(2, 0.5)
)
# Define spatial decay prior using spT.decay
decay_prior <- spTimer::spT.decay(
  distribution = spTimer::Gamm(a = 2, b = 1),  # Gamma prior for phi
  tuning = 0.1,  # Tuning parameter 
  npoints = NULL,  # Number of points for discretization
  value = NULL     # Fixed value for phi 
)

# Fit the model
model <- spTimer::spT.Gibbs(
  formula = formula,
  data = data,
  coords = coords,
  model = "GP",
  priors = priors,
  nItr = 100,
  nBurn = 10,
  cov.fnc = "exponential",
  tol.dist = 1e-3
)

# Check model summary
summary(model)
```

```{r}
# Load the dataset
data <- read.csv("/Users/varshap/Downloads/ST5188/Data/Final/TT Split/changi_train_wide.csv")

# Randomly select 1000 rows
set.seed(123)  # Ensures reproducibility
data_sample <- data %>% sample_n(1000)

# Add small random noise to coordinates
data_sample <- data_sample %>%
  mutate(
    x = x + runif(n(), -0.1, 0.1),  # Add noise to x
    y = y + runif(n(), -0.1, 0.1)   # Add noise to y
  )

# Rescale coordinates
# data_sample <- data_sample %>%
#   mutate(
#     x = (x - min(x)) / (max(x) - min(x)),
#     y = (y - min(y)) / (max(y) - min(y))
#   )

# Extract the month columns 
month_cols <- names(data_sample)[grepl("\\d{4}", names(data_sample))]  # Select columns containing years

# Extract coordinates
coords <- as.matrix(data_sample[, c("x", "y")])

# Define priors
priors <- spTimer::spT.priors(
  model = "GP",
  inv.var.prior = spTimer::Gamm(2, 0.1),
  beta.prior = spTimer::Norm(0, 10^2)
)

priors <- spTimer::spT.priors(
  model = "GP",
  inv.var.prior = spTimer::Gamm(2, 0.1),  # Adjust the shape and scale
  beta.prior = spTimer::Norm(0, 10^2) #,     # Increase variance
  #phi.prior = spTimer::Gamm(2, 0.5)       # controls spatial decay
)

# Define spatial decay prior using spT.decay
decay_prior <- spTimer::spT.decay(
  distribution = spTimer::Gamm(a = 2, b = 1),  # Gamma prior for phi
  tuning = 0.1,  # Tuning parameter 
  npoints = NULL,  # Number of points for discretization
  value = NULL     # Fixed value for phi 
)

# Loop over all months and fit the model
models <- list()
for (month in month_cols) {
  formula <- as.formula(paste0("`", month, "` ~ x + y"))  # Create formula dynamically
  
  # Fit the model
  model <- spTimer::spT.Gibbs(
    formula = formula,
    data = data_sample,
    coords = coords,
    model = "GP",
    priors = priors,
    nItr = 100,
    nBurn = 10,
    cov.fnc = "exponential",
    tol.dist = 1e-2
  )
  
  # Store model in the list
  models[[month]] <- model
  
  # Print progress
  cat("Finished model for:", month, "\n")
}

# Print summaries of all models
lapply(models, summary)

```

# Load the dataset
data <- read.csv("/Users/varshap/Downloads/ST5188/Data/Final/TT Split/changi_train_wide.csv")

# Randomly select 1000 rows
set.seed(123)  # Ensures reproducibility
data_sample <- data %>% sample_n(1000)

# Add small random noise to coordinates
data_sample <- data_sample %>%
  mutate(
    x = x + runif(n(), -0.1, 0.1),  # Add noise to x
    y = y + runif(n(), -0.1, 0.1)   # Add noise to y
  )

# Rescale coordinates
# data_sample <- data_sample %>%
#   mutate(
#     x = (x - min(x)) / (max(x) - min(x)),
#     y = (y - min(y)) / (max(y) - min(y))
#   )

# Extract the month columns (assuming they are named like 'Mar.Apr.2000', 'May.Jun.2000', etc.)
month_cols <- names(data_sample)[grepl("\\d{4}", names(data_sample))]  # Select columns containing years

# Extract coordinates
coords <- as.matrix(data_sample[, c("x", "y")])

# Define priors
priors <- spTimer::spT.priors(
  model = "GP",
  inv.var.prior = spTimer::Gamm(2, 0.1),
  beta.prior = spTimer::Norm(0, 10^2)
)

priors <- spTimer::spT.priors(
  model = "GP",
  inv.var.prior = spTimer::Gamm(2, 0.1),  # Adjust the shape and scale
  beta.prior = spTimer::Norm(0, 10^2) #,     # Increase variance
  #phi.prior = spTimer::Gamm(2, 0.5)       # controls spatial decay
)

# Define spatial decay prior using spT.decay
decay_prior <- spTimer::spT.decay(
  distribution = spTimer::Gamm(a = 2, b = 1),  # Gamma prior for phi
  tuning = 0.1,  # Tuning parameter 
  npoints = NULL,  # Number of points for discretization
  value = NULL     # Fixed value for phi 
)

# Loop over all months and fit the model
models <- list()
for (month in month_cols) {
  formula <- as.formula(paste0("`", month, "` ~ x + y"))  # Create formula dynamically
  
  # Fit the model
  model <- spTimer::spT.Gibbs(
    formula = formula,
    data = data_sample,
    coords = coords,
    model = "GP",
    priors = priors,
    decay=decay_prior,
    nItr = 100,
    nBurn = 10,
    cov.fnc = "exponential",
    tol.dist = 1e-2
  )
  
  # Store model in the list
  models[[month]] <- model
  
  # Print progress
  cat("Finished model for:", month, "\n")
}

# Print summaries of all models
lapply(models, summary)

